\section{BPMN Sample Catalog}
\label{sec:BpmnSamples}

The repository contains a growing set of BPMN models under
\filepath{src/test/resources/bpmn}. They serve two purposes:
\begin{itemize}
\item they are regression inputs for the scaffolder and its tests,
\item they are concrete learning models for understanding one BPMN feature at a time.
\end{itemize}

The samples are intentionally small. Most of them isolate one feature rather than trying
to model a realistic end-to-end business process in full detail.

\begin{longtable}{|>{\raggedright\arraybackslash}p{0.38\textwidth}|>{\raggedright\arraybackslash}p{0.54\textwidth}|}
\hline
\rowcolor[gray]{0.9}\textbf{Model} & \textbf{Primary focus} \\
\hline
\endfirsthead
\hline
\rowcolor[gray]{0.9}\textbf{Model} & \textbf{Primary focus} \\
\hline
\endhead
\filepath{invoice_receipt.bpmn} &
Baseline process with start, service work, review, approval or rejection, and terminal notification. This is the best first model to inspect. \\
\hline
\filepath{order_fulfillment.bpmn} &
Legacy process sample kept as an additional reference model. It is useful for checking that generator changes are not too tightly coupled to the invoice-oriented examples. \\
\hline
\filepath{invoice_receipt_reminder.bpmn} &
Intermediate timer catch event between normal process steps. Demonstrates delay-driven routing without a boundary event. \\
\hline
\filepath{invoice_message_exchange.bpmn} &
Intermediate message throw and catch events. Shows explicit Kafka topic mappings for message-based collaboration. \\
\hline
\filepath{invoice_signal_exchange.bpmn} &
Intermediate signal throw and catch events. Useful for contrasting signal semantics with the more direct message-event mapping. \\
\hline
\filepath{invoice_review_deadline.bpmn} &
Interrupting timer boundary on a task. Demonstrates timeout handling and normal-flow suppression after deadline expiry. \\
\hline
\filepath{invoice_review_reminder_non_interrupting.bpmn} &
Non-interrupting timer boundary on a task. Demonstrates reminder-style branching without cancelling the main task. \\
\hline
\filepath{invoice_processing_error.bpmn} &
Interrupting error boundary on a task, driven by task failure events from the generated runtime. \\
\hline
\filepath{invoice_review_escalation.bpmn} &
Interrupting escalation boundary on a task. Shows how escalation differs from failure as an explicit lifecycle outcome. \\
\hline
\filepath{invoice_call_activity.bpmn} &
Call-activity request and reply mapping. Demonstrates the generated call and reply topics plus manual completion of the invoked activity. \\
\hline
\filepath{invoice_review_subprocess.bpmn} &
Embedded subprocess with explicit entry and completion handlers. This is the basic scope-aware subprocess sample. \\
\hline
\filepath{invoice_nested_subprocess.bpmn} &
Nested embedded subprocesses. Useful for understanding recursive scope generation and subprocess-specific input and output channels. \\
\hline
\filepath{invoice_subprocess_deadline.bpmn} &
Interrupting timer boundary attached to a subprocess scope. Demonstrates cancellation at the subprocess level rather than at one task. \\
\hline
\filepath{invoice_subprocess_reminder_non_interrupting.bpmn} &
Non-interrupting timer boundary attached to a subprocess scope. Shows branch-without-cancel behavior for a whole scope. \\
\hline
\filepath{invoice_subprocess_error.bpmn} &
Interrupting error boundary attached to a subprocess scope. Demonstrates subprocess-level failure propagation. \\
\hline
\filepath{invoice_event_subprocess_message.bpmn} &
Non-interrupting embedded event subprocess with message start. Shows side-flow activation within an enclosing subprocess scope. \\
\hline
\filepath{invoice_event_subprocess_interrupting_message.bpmn} &
Interrupting embedded event subprocess with message start. Demonstrates cancellation of the enclosing scope before the event subprocess branch starts. \\
\hline
\filepath{invoice_event_subprocess_timer.bpmn} &
Embedded event subprocess with timer start. Demonstrates scoped delayed activation tied to the enclosing subprocess lifecycle. \\
\hline
\filepath{invoice_event_subprocess_error.bpmn} &
Embedded event subprocess with error start. Demonstrates start-on-failure semantics driven by the canonical lifecycle stream. \\
\hline
\filepath{invoice_event_subprocess_escalation.bpmn} &
Embedded event subprocess with escalation start. Demonstrates start-on-escalation semantics within a subprocess scope. \\
\hline
\filepath{data_pipeline_demo.bpmn} &
Minimal data pipeline with three plugin-annotated tasks: JSON Transform, Field Filter, and KV Enricher. The simplest plugin catalog example. \\
\hline
\filepath{order_events_pipeline.bpmn} &
Realistic order event pipeline with 8 plugin tasks (transform, coerce, filter, enrich, validate, UUID-inject, timestamp-normalize, mask) and an XOR gateway routing by amount using runtime-evaluated conditions. Designed to be used with \texttt{--connect} for source/sink generation. \\
\hline
\filepath{log_processing_pipeline.bpmn} &
Log processing pipeline demonstrating regex extraction, type coercion, string templating, JSON flattening, schema validation, and explicit field masking in a single linear flow. Also designed for \texttt{--connect} source/sink integration. \\
\hline
\filepath{custom_activity_demo.bpmn} &
Custom activity round-trip example. A ServiceTask annotated with \texttt{camunda:plugin="custom"} triggers contract interface generation and delegating worker creation for hand-written implementations. \\
\hline
\end{longtable}

Taken together, the sample set functions as a compact executable coverage matrix. When a new BPMN function
is added to the generator, adding a dedicated sample model alongside it has become the preferred way to
document the feature, test the generator, and provide a runnable teaching example.
