\section{Code-Generation Targets}
\label{sec:RustTarget}

The scaffolder is target-parameterised: from a single BPMN model it can render a
generated project in more than one implementation language. The mature target is
Java (Quarkus / SmallRye Reactive Messaging); a Rust target is in progress. Both
targets emit workers that speak the same Kafka wire protocol, so a single process
can, in principle, be realised as a mix of Java and Rust workers, and the same
Java monitor observes either.

\subsection{Template organisation}

Template groups are separated by target under \filepath{durga-tools/src/main/resources/}:

\begin{itemize}
\item \filepath{templates-java/} --- the Java target (\filepath{scaffold.stg} and
  companions). This is the default and the only fully implemented target.
\item \filepath{templates-rust/} --- the Rust target (in progress).
\end{itemize}

The attribute model passed from the generator (process id, task id, plugin
reference, plugin configuration, channels, lineage) is identical across targets;
only the rendered syntax differs. Loading the template group is the single
switch point in \texttt{BpmnScaffolder}.

\subsection{Shared source of truth}

To keep the targets in lock-step, the following remain single sources of truth
consumed by both:

\begin{itemize}
\item \textbf{Plugin catalog} --- the YAML descriptors under \filepath{plugins/}
  define plugin identity, category, and the typed configuration schema. A model
  scaffolds identically for either target; the same plugin configuration string
  (e.g.\ \texttt{id:order\_id, customer.email:customer\_email}) is honoured by the
  Java and Rust plugin implementations alike.
\item \textbf{Lifecycle wire format} --- the \texttt{ProcessEvent} JSON (camelCase
  fields, SCREAMING\_SNAKE event types and statuses) is byte-compatible across
  targets, so the monitoring topology, alarm model, and dashboard treat
  Rust-generated tasks exactly as Java tasks.
\item \textbf{Chaining topics} --- both targets use identical topic names: a task
  consumes its predecessor's \texttt{\%processId\%\_\%predecessor\%\_output}, a routing
  gateway writes to each target's \texttt{\%processId\%\_\%target\%\_input}, and the
  first task consumes \texttt{\%processId\%\_start}. A Java and a Rust realisation of the
  same model are therefore wire-interchangeable on every topic --- a mixed fleet works,
  and a shadow of either target can validate a production of either target.
\end{itemize}

\subsection{The Rust plugin runtime (\texttt{durga-rust})}

The \filepath{durga-rust/} Cargo crate is the Rust counterpart to the Java
\filepath{durga-runtime} plugin layer. Generated Rust workers depend on it
directly --- there are no adapters onto the Java plugins; each plugin is a native
Rust implementation.

The \texttt{Plugin} trait mirrors the Java \texttt{Plugin} interface, including its
dispatch defaults:

\begin{itemize}
\item \texttt{execute\_bytes(payload, config)} --- binary entry point; the default
  UTF-8 decodes and delegates to \texttt{execute\_str}, then re-encodes. Binary
  plugins override this directly.
\item \texttt{execute\_str(payload, config)} --- text / JSON entry point; text
  plugins override this. The default is unsupported, matching the Java default.
\item \texttt{execute\_with\_result(payload, config, context)} --- the structured
  entry point the generated worker calls, carrying a \texttt{PluginExecutionContext};
  the default wraps \texttt{execute\_bytes} with a content-based idempotency key and the
  \texttt{PAYLOAD} disposition, ignoring the context. Side-effecting plugins override it
  to suppress effects when \texttt{context.validation\_mode()} is true.
\item \texttt{idempotency\_key(payload, config)} --- SHA-256 of payload then
  config, byte-identical to the Java default.
\end{itemize}

\texttt{PluginResult} carries the same \texttt{OutputDisposition}
(\texttt{PAYLOAD}, \texttt{PASSTHROUGH}, \texttt{SIDE\_EFFECT}) and
\texttt{ErrorStrategy} (\texttt{FAIL}, \texttt{SKIP}, \texttt{DLQ}) semantics as
the Java record, so routing and inspection plugins declare \texttt{PASSTHROUGH}
and the generated worker preserves the payload identically on either target.

\subsection{Validation shadows}

Validation mode (Section~\ref{sec:ValidationMode}) is emitted on both targets. With
\texttt{--validation} the scaffolder generates the \emph{whole} process again, wired to
read the live production topics through a dedicated consumer group, suppress side effects
(via the validation \texttt{PluginExecutionContext}), divert each task's output to a
\texttt{-validation} topic, and emit lifecycle events to the per-process
\texttt{process-events-\%processId\%-validation} topic. There is no separate shadow-worker class or validation
binary and no separate candidate wire record: both targets emit ordinary
\texttt{ProcessEvent} records, and because their chaining topics are identical, a shadow
of either target validates a production of either target.

\subsection{Status}

Implemented so far: the crate contracts (\texttt{Plugin}, \texttt{PluginResult},
\texttt{PluginExecutionContext}, \texttt{ProcessEvent}, dot-path field helpers), the
pipeline plugin ports, the pure worker decision logic (\texttt{plan\_worker\_output}),
the Rust worker / gateway and project (\texttt{Cargo.toml}, \texttt{lib.rs},
\texttt{build.rs}) templates, and the \texttt{-{}-target rust} switch in the CLI. Rust
workers use a mature Kafka client (\texttt{rdkafka}) and an async runtime
(\texttt{tokio}) for the same consume--transform--emit loop the plain-Kafka Java
worker performs. Rust remains the less-mature target: it currently covers plugin
tasks, custom-task stubs (with contract modules and \texttt{build.rs} entries),
exclusive gateways, and validation mode in a start/tasks/gateways/end topology, and
reports and skips unsupported constructs.
