\section{Configuration And Operations}
\label{sec:ConfigurationOperations}

\subsection{Scaffolder invocation}

The repository includes a convenience script at \filepath{durga}:

\begin{verbatim}
./durga path/to/process.bpmn [flags...]
\end{verbatim}

It auto-builds the JAR if missing and passes all arguments through to the
scaffolder. Equivalent to:

\begin{verbatim}
mvn -q package -DskipTests
java -jar target/durga-0.1.0-beta.1.jar path/to/process.bpmn [flags...]
\end{verbatim}

Available flags: \texttt{--dry-run}, \texttt{--out <dir>},
\texttt{--transactions}, \texttt{--separate-workers}, \texttt{--strimzi},
\texttt{--connect}, \texttt{--validation}, \texttt{--target java|rust},
\texttt{--event-topic <topic>}, \texttt{--process-id <id>},
\texttt{--package <pkg>}, \texttt{--retention <hours>}.

\texttt{--validation} additionally generates a validation-mode shadow worker for
each plugin task (Java or Rust), so a candidate implementation can be run against
real input on a dedicated consumer group with side effects suppressed and its
output compared against production. It also registers the \\
\texttt{validation-candidate-outputs} and \texttt{validation-results} topics. See
Section~\ref{sec:ValidationMode}. The default output (without the flag) is
unchanged.

\texttt{--retention <hours>} sets Kafka topic retention in hours
(defaults to 168\,h/7\,days with a warning). Use \texttt{-1} for
infinite retention. The canonical event topic (configured via
\texttt{--event-topic}) always has infinite retention; the
\texttt{\%process\%\_state} topic always uses compaction. Wired
into both \texttt{topics.sh} and Strimzi \texttt{topics.yml}.
Accepts \texttt{h}, \texttt{d}, \texttt{w}, \texttt{m} suffixes
for hours, days, weeks, or 30-day months.

For production, do not accept generated retention defaults without review.
Retention is part of the process contract: it controls replay windows,
diagnostic history, storage cost and monitoring recovery. Infinite retention
requires an explicit storage and compaction strategy.

\texttt{--package <pkg>} sets the Java package for generated code
(defaults to \texttt{org.gautelis.durga.\\generated} with a warning).
Operational classes are placed directly in the package; probe and
exerciser classes in \texttt{<pkg>.probes}.

\texttt{--event-topic <topic>} overrides the physical Kafka topic name for
canonical lifecycle events. The default is \texttt{process-events-<processId>},
so each pipeline writes to its own topic by default. Use this flag to direct
multiple pipelines to a shared topic (e.g. \texttt{--event-topic process-events})
or to use a custom naming scheme. The logical channel names
\texttt{process-events} and \texttt{process-events-monitor} in the
generated code stay unchanged; only the physical topic mapping in
\texttt{application.yml} is affected.

\texttt{--process-id <id>} overrides the process identifier read from the
BPMN file. This is the name used in topic names, consumer group IDs,
generated class names, and the \texttt{durga-<id>} consumer group prefix
for the \texttt{ProcessStateStore}. Pinning the process ID keeps Kafka
group coordination stable if the BPMN tool regenerates with different
element identifiers.

\subsection{Local runtime baseline}

The normal local environment consists of:
\begin{itemize}
\item the bundled Kafka setup under \filepath{setup/},
\item generated topic configuration in the scaffolded project,
\item the optional monitoring app and dashboard.
\end{itemize}

The repository is intentionally optimized for a local single-broker loop so the generated
processes and monitoring projections can be explored quickly. This local
baseline is not a production topology; production deployments need broker
replication, authenticated clients, managed secrets, resource limits, topic
ownership and backup/replay policy.

\subsection{Important configuration locations}

The main configuration locations are:
\begin{itemize}
\item repository-level runtime defaults in \filepath{src/main/resources/application.yml},
\item generated-project channel configuration in the scaffolded \texttt{application.yml},
\item local Kafka infrastructure in \filepath{setup/docker-compose.yml},
\item generated topic setup scripts in scaffolded output directories.
\end{itemize}

\subsection{Operational expectations}

In normal local startup:
\begin{itemize}
\item Kafka UI may log temporary connection failures before the broker is ready,
\item generated runtimes rely on explicit topics and channels rather than implicit discovery,
\item the monitoring app may report a transitional Streams state before its stores become queryable.
\end{itemize}

\subsection{Where state lives}

Durga uses several layers of state:
\begin{itemize}
\item transport history on execution topics,
\item canonical lifecycle history on \texttt{process-events},
\item materialized latest state and counts in Kafka Streams stores and corresponding topics,
\item local waiting and cancellation bookkeeping inside some generated handlers.
\end{itemize}

Operationally, the first three layers are Kafka-backed and can be recovered or
replayed within their configured retention windows. The last layer is process
memory and should be treated as transient unless the generated handler explicitly
persists its coordination state.

\subsection{Recommended local development loop}

For practical local work:
\begin{enumerate}
\item start Kafka,
\item build the repository or generated project,
\item create topics,
\item run the generated process or a demo publisher,
\item start the monitoring app if query visibility is needed,
\item inspect lifecycle events, counts and latency.
\end{enumerate}
