\section{Kafka Connect Integration}
\label{sec:Connect}

Kafka Connect handles edge I/O. Durga owns the process and topic topology;
Connect owns the movement between Kafka and external systems. The scaffolder's
\texttt{--connect} flag generates connector skeletons for the pipeline's
generic intake/output topics and, when BPMN data stores are present, for
data-store-specific source and sink edges.

\subsection{Architecture}

\begin{figure}[!ht]
    \centering
    \begin{tikzpicture}[node distance=16mm,>=latex]
        \tikzstyle{box}=[draw=iptoMuted, rounded corners=2pt, line width=0.8pt, align=center, minimum width=18mm, minimum height=10mm, fill=white]
        \tikzstyle{accentbox}=[box, fill=iptoAccent!12]
        \tikzstyle{ext}=[draw=iptoMuted!50, rounded corners=2pt, line width=0.6pt, align=center, minimum width=22mm, minimum height=9mm, fill=white, dashed]

        \node[ext] (extsrc) {External\\system source};
        \node[box, right=5mm of extsrc] (connsrc) {Connect\\source};
        \node[box, right=5mm of connsrc] (intake) {Intake\\topic};
        \node[box, right=5mm of intake] (pipeline) {Durga\\pipeline};
        \node[box, right=5mm of pipeline] (output) {Output\\topic};
        \node[box, right=5mm of output] (connsink) {Connect\\sink};
        \node[ext, right=5mm of connsink] (extsink) {External\\system sink};

        \node[box, below=12mm of connsrc] (msgsrc) {Connect\\source};
        \node[ext, left=5mm of msgsrc] (extmsg) {External\\message};
        \node[box, right=5mm of msgsrc] (msgtopic) {Message\\topic};
        \node[accentbox, right=5mm of msgtopic] (pipe) {Pipeline\\receive};

        \draw[->, thick] (extsrc) -- (connsrc);
        \draw[->, thick] (connsrc) -- (intake);
        \draw[->, thick] (intake) -- (pipeline);
        \draw[->, thick] (pipeline) -- (output);
        \draw[->, thick] (output) -- (connsink);
        \draw[->, thick] (connsink) -- (extsink);

        \draw[->, thick] (extmsg) -- (msgsrc);
        \draw[->, thick] (msgsrc) -- (msgtopic);
        \draw[->, thick] (msgtopic) -- (pipe);
    \end{tikzpicture}
    \caption{Kafka Connect integration: sources feed intake topics, sinks consume output topics}
    \label{fig:DurgaConnect}
\end{figure}

\subsection{Enabling}

\begin{verbatim}
./durga path/to/process.bpmn --connect
\end{verbatim}

This adds a \filepath{connect/} directory to the generated project:

\begin{verbatim}
connect/connect-source.json
connect/connect-sink.json
connect/data-stores/*.json
connect.sh
\end{verbatim}

\subsection{Source connectors}

\texttt{connect-source.json} contains connector configuration for:
\begin{itemize}
\item the process start topic (\texttt{\%processId\%\_start}),
\item message catch and receive task topics.
\end{itemize}

The user fills in the connector class and system-specific settings:

\begin{verbatim}
{
  "name": "invoice_receipt-source",
  "config": {
    "connector.class": "...",
    "tasks.max": "1",
    "invoice_receipt_start": {
      "kafka.topic": "invoice_receipt_start"
    }
  }
}
\end{verbatim}

\subsection{Sink connectors}

\texttt{connect-sink.json} contains connector configuration for:
\begin{itemize}
\item terminal task outputs,
\item the canonical \texttt{process-events-\{processId\}} topic (for auditing or replication),
\item message throw and send task topics.
\end{itemize}

\subsection{Data-store connectors}

When BPMN data stores are connected to tasks by data associations, Durga also
generates one connector skeleton per data-store edge:

\begin{itemize}
\item a data store used as a task input source becomes a source connector
      skeleton that writes to the corresponding task input topic,
\item a data store used as a task output target becomes a sink connector
      skeleton that reads from the corresponding task output topic.
\end{itemize}

For example, a task \texttt{enrich\_data} that writes to
\texttt{s3\_warehouse}, \texttt{neo4j\_customer\_graph}, and
\texttt{postgresql\_customer\_store} produces files such as:

\begin{verbatim}
connect/data-stores/data_pipeline_demo-s3-warehouse-sink.json
connect/data-stores/data_pipeline_demo-neo4j-customer-graph-sink.json
connect/data-stores/data_pipeline_demo-postgresql-customer-store-sink.json
\end{verbatim}

These skeletons consume the existing Durga task output topic, for example
\texttt{data\_pipeline\_demo\\ \_enrich\_data\_output}. The BPMN data object
remains a logical asset; it is not converted into a Kafka topic.

\subsection{Deploying}

The generated \texttt{connect.sh} posts all connector JSON files under
\filepath{connect/} and \filepath{connect/data-stores/} to the Connect REST API:

\begin{verbatim}
CONNECT_URL=http://kafka-connect:8083 ./connect.sh
\end{verbatim}

Connector status can be checked with:

\begin{verbatim}
curl http://kafka-connect:8083/connectors/invoice_receipt-source/status
curl http://kafka-connect:8083/connectors/invoice_receipt-sink/status
curl http://kafka-connect:8083/connectors/data_pipeline_demo-s3-warehouse-sink/status
\end{verbatim}

\subsection{Common connector classes}

\begin{longtable}{|>{\raggedright\arraybackslash}p{0.42\textwidth}|>{\raggedright\arraybackslash}p{0.50\textwidth}|}
\hline
\rowcolor[gray]{0.9}\textbf{Connector} & \textbf{Use case} \\
\hline
\texttt{DebeziumSourceConnector} & CDC from PostgreSQL, MySQL, MongoDB \\
\texttt{JdbcSourceConnector} & Polling queries from relational databases \\
\texttt{S3SinkConnector} & Flush to S3-compatible object stores \\
\texttt{ElasticsearchSinkConnector} & Index to Elasticsearch \\
\texttt{HttpSinkConnector} & POST to REST APIs \\
\texttt{JdbcSinkConnector} & Write to relational databases \\
\texttt{Neo4jSinkConnector} & Write graph changes to Neo4j \\
\texttt{BigQuerySinkConnector} & Load into Google BigQuery \\
\hline
\end{longtable}

\subsection{Limitations}

\begin{itemize}
\item Connector configs are generated as stubs. The user must fill in
      connector-specific settings (connection URLs, credentials, table names,
      bucket names, converters and transforms).
\item Connector class names for data stores are hints. The deployed Connect
      cluster must actually have the selected connector plugin installed.
\item The scaffolder does not validate connector configs against a live
      Connect instance.
\item Single Message Transform (SMT) chains are not generated. The user
      adds transforms in the connector config as needed.
\end{itemize}
