\section{Data Assets}
\label{sec:DataAssets}

Durga treats BPMN data objects as logical pipeline assets, not as Kafka
topics. Sequence flows still drive execution and Kafka still carries process
events, worker input/output messages, BPMN messages and BPMN signals. Data
objects describe the business artifacts that tasks read and write.

\subsection{Model}

A BPMN data object represents a named dataset or artifact, for example
\texttt{RawCustomerData}, \texttt{NormalizedCustomerData}, or
\texttt{CustomerGraphDelta}. A BPMN data store represents a physical source or
target such as S3, PostgreSQL, Neo4j, a local file store, or an API.

Data input and output associations connect executable activities to those
assets:

\begin{verbatim}
Transform Data
  input:  RawCustomerData
  output: NormalizedCustomerData

Enrich Data
  input:  FilteredCustomerData
  output: EnrichedCustomerData
  stores: S3Warehouse, Neo4jCustomerGraph, PostgresCustomerStore
\end{verbatim}

The scaffolder collects these declarations into \filepath{summary.json} and
the generated README. The declarations are metadata and contracts; they do not
change routing by themselves.

\subsection{Runtime handles}

Large datasets should normally travel by reference rather than as Kafka message
payload bytes. The generated runtime includes \texttt{DataHandle}:

\begin{verbatim}
DataHandle(
    name,
    uri,
    mediaType,
    schema,
    metadata)
\end{verbatim}

A process event payload can carry a serialized data handle such as an S3 URI,
media type, schema identifier, row count, checksum, or partition metadata. This
keeps lifecycle events small while preserving enough information for monitoring
and lineage.

Durga now includes experimental store plugins for this pattern. The
\texttt{object-store-collector} writes the current payload to a local file-backed
object store and forwards a \texttt{DataHandle}-style reference. The
\texttt{object-store-extractor} resolves that reference and emits the referenced
bytes to the next topic. Generated plugin and custom workers can also use
\texttt{handleMode=manual} to pass handles through as plugin payloads, or
\texttt{handleMode=materialize} to load the referenced bytes, run the plugin, and
write the modified result back to the object store as a new handle. The current
implementation is local-file backed; S3, GCS, Azure Blob, and other object-store
adapters remain production hardening work.

\subsection{Extension properties}

Data objects and data stores may use Camunda extension properties for Durga
metadata:

\begin{verbatim}
schema    = schemas/enriched-customer-data.schema.json
mediaType = application/json
kind      = s3
uri       = s3://warehouse/customer/enriched/
\end{verbatim}

For data stores, \texttt{kind} identifies the adapter family. Current examples
include \texttt{s3}, \texttt{postgresql}, and \texttt{neo4j}. If \texttt{kind}
is omitted, later tooling may infer it from the URI scheme.

\subsection{Lineage and metadata events}

The implementation records data objects, data stores, and task associations in
generated metadata. Generated plugin and custom workers also emit a
Vannak-compatible \texttt{DataIndividualMetadataEvent} companion record to the
\texttt{vannak-metadata-events} Kafka topic after successful execution. That
event captures process/activity context, payload size and checksums, plugin
identity/configuration, output fields, and \texttt{DataHandle}/format metadata
when present.

The remaining direction is to bind BPMN data associations more tightly to those
events, so task completions can report:

\begin{verbatim}
read:   [RawCustomerData]
wrote:  [EnrichedCustomerData]
stores: [S3Warehouse, Neo4jCustomerGraph]
\end{verbatim}

This keeps the process engine focused on orchestration while giving pipeline
operators the data-plane visibility needed for real workloads.
