\section{Deployment}
\label{sec:Deployment}

Durga-generated projects are plain Java applications that connect to Kafka.
They can be deployed as JARs, Docker containers, or Kubernetes workloads
with no Durga-specific runtime service beyond the generated code and a Kafka
cluster.

\subsection{Architecture}

Each generated process instance is a set of Quarkus-based workers (one per
BPMN task) communicating through Kafka topics. There is no central orchestrator
at runtime --- workers consume from input topics and produce to output topics
independently. This means:

\begin{itemize}
\item every worker is horizontally scalable via Kafka consumer groups,
\item worker deployment is stateless; durable state lives in Kafka topics and
  Kafka Streams stores, while some generated handlers still keep transient
  in-memory coordination state,
\item the generated shaded JAR is self-contained (all dependencies bundled).
\end{itemize}

\subsection{Local development \& testing}

The recommended local loop uses Docker Compose for Kafka and direct JVM
execution for the pipeline:

\begin{verbatim}
# Start Kafka
cd setup && docker compose up -d

# Build and run the pipeline
cd generated/<project>
START_KAFKA=false BOOTSTRAP=localhost:9094 ./run-local.sh
\end{verbatim}

For iterative development, keep Kafka running and rebuild the pipeline:

\begin{verbatim}
mvn -q clean package -DskipTests
java -cp target/<project>-all.jar \
  org.gautelis.durga.generated.probes.<Process>Starter localhost:9094
\end{verbatim}

To test with the monitoring dashboard:

\begin{enumerate}
\item Start Kafka as above.
\item Start the monitoring app:
  \begin{verbatim}
  cd setup
  START_KAFKA=false ./demo-monitoring.sh
  \end{verbatim}
\item Run pipeline scenarios via the generated helper scripts.
\item Open \texttt{http://localhost:18081} for the Svelte dashboard
      or \\ \texttt{http://localhost:18081/api} for the REST API.
\end{enumerate}

\subsection{Connection to Kafka}

Kafka bootstrap servers are resolved slightly differently depending on the
process, but all paths end at the same hardcoded default \texttt{localhost:9094}:

\begin{enumerate}
\item The monitoring app takes an optional first CLI argument, then falls back to
the JVM system property \texttt{kafka.bootstrap.servers}, then the default.
\item Generated standalone (non-Quarkus) workers read the JVM system property
\texttt{kafka.bootstrap.servers}, then the default.
\item Generated Quarkus workers read the environment variable
\texttt{KAFKA\_BOOTSTRAP\_SERVERS} (via \texttt{application.yml}), then the default.
\end{enumerate}

In containers the generated \texttt{Dockerfile} bridges the environment variable
into the JVM property, so setting \texttt{KAFKA\_BOOTSTRAP\_SERVERS} works uniformly
across worker styles (see below).

In Docker and Kubernetes, set \texttt{KAFKA\_BOOTSTRAP\_SERVERS} as a container
environment variable. The generated \texttt{Dockerfile} passes it to the JVM:

\begin{verbatim}
ENTRYPOINT ["java", "-Xms64m", "-Xmx256m", \
  "-Dkafka.bootstrap.servers=${KAFKA_BOOTSTRAP_SERVERS:-localhost:9094}", \
  "-jar", "/app/pipeline.jar"]
\end{verbatim}

\subsection{Topic provisioning}

Kafka topics must exist before the pipeline starts. There are three approaches,
each supported by the generated artifacts.

\subsubsection{Pre-creation via shell script}

The generated \texttt{topics.sh} creates all required topics using the Kafka
CLI tools bundled with the Kafka distribution:

\begin{verbatim}
KAFKA_BOOTSTRAP_SERVERS=kafka-broker:9092 ./topics.sh
\end{verbatim}

This is suitable for development, staging, and controlled manual production
setups. For production, review partition count, replication factor, cleanup
policy and retention before running the script. Automated environments should
prefer an init job, a CI/CD provisioning step, or declarative topic management
that is owned by the platform team.

\subsubsection{Init container}

For Kubernetes deployments where the Kafka broker shares the cluster,
add an init container to the generated \texttt{k8s.yml}:

\begin{verbatim}
initContainers:
  - name: topic-init
    image: bitnami/kafka:latest
    command: ['/bin/sh', '-c']
    args:
      - |
        /opt/bitnami/kafka/bin/kafka-topics.sh --create \
          --bootstrap-server $KAFKA_BOOTSTRAP_SERVERS \
          --topic <name> --partitions 3 --replication-factor 1
\end{verbatim}

\subsubsection{Kafka operator (Strimzi)}

In Kubernetes environments running the Strimzi operator, topics can be
declared as \texttt{KafkaTopic} custom resources. A generated
\texttt{kafka-topics.yml} CRD manifest can be applied alongside the
Deployment:

\begin{verbatim}
apiVersion: kafka.strimzi.io/v1beta2
kind: KafkaTopic
metadata:
  name: <process>_<task>_input
  labels:
    strimzi.io/cluster: my-kafka
spec:
  partitions: 3
  replicas: 1
---
\end{verbatim}

The scaffolder can generate Strimzi-oriented output when requested. Treat the
generated manifest as a starting point: set replicas, partitions, retention and
labels to match the target Kafka cluster policy before applying it.

\subsection{Test environment deployment}

A Docker Compose-based test environment is the fastest way to validate a
pipeline end-to-end. The repository's \texttt{setup/docker-compose.yml}
provides a single-node Kafka broker and Kafka UI.

Add a \texttt{docker-compose.test.yml} to the generated project:

\begin{verbatim}
version: "3"
services:
  pipeline:
    build: .
    environment:
      KAFKA_BOOTSTRAP_SERVERS: kafka:9092
    depends_on:
      - kafka
  kafka:
    image: bitnami/kafka:latest
    environment:
      KAFKA_CFG_NODE_ID: 1
      KAFKA_CFG_PROCESS_ROLES: broker,controller
      KAFKA_CFG_LISTENERS: PLAINTEXT://:9092,CONTROLLER://:9093
      KAFKA_CFG_ADVERTISED_LISTENERS: PLAINTEXT://kafka:9092
      KAFKA_CFG_CONTROLLER_LISTENER_NAMES: CONTROLLER
      KAFKA_CFG_CONTROLLER_QUORUM_VOTERS: 1@kafka:9093
\end{verbatim}

Run with:
\begin{verbatim}
docker compose -f docker-compose.test.yml up --build
\end{verbatim}

\subsection{Production deployment}

The generated \texttt{deploy.sh} supports three targets:

\begin{description}
\item[JAR] \texttt{./deploy.sh} --- builds the shaded JAR and runs locally.
  Suitable for development, VMs and bare-metal deployments when process
  supervision is provided externally.
\item[Docker] \texttt{DOCKER=true ./deploy.sh} --- builds a Docker image
  tagged \texttt{<processId>:latest}.
\item[Kubernetes] \texttt{DOCKER=true K8S=true ./deploy.sh} --- builds the
  image and applies the generated \texttt{k8s.yml}.
\end{description}

\subsubsection{Docker image}

The generated \texttt{Dockerfile} uses \texttt{eclipse-temurin:21-jre-alpine},
runs as a non-root user, and includes a process-based health check. In
production, replace or supplement process checks with application-level
readiness that verifies Kafka connectivity and worker readiness:

\begin{verbatim}
FROM eclipse-temurin:21-jre-alpine
RUN addgroup -S durga && adduser -S durga -G durga
USER durga
COPY target/*-all.jar /app/pipeline.jar
HEALTHCHECK --interval=30s ...
ENTRYPOINT ["java", "-Xms64m", "-Xmx256m", \
  "-Dkafka.bootstrap.servers=${KAFKA_BOOTSTRAP_SERVERS:-...}", \
  "-jar", "/app/pipeline.jar"]
\end{verbatim}

\subsubsection{Kubernetes}

The generated \texttt{k8s.yml} provides:

\begin{itemize}
\item A \texttt{Deployment} with 2 replicas (configurable), baseline resource
  requests and limits,
\item A \texttt{Service} exposing port 8080,
\item Exec-based liveness and readiness probes.
\end{itemize}

\subsection{Scaling}

Workers scale horizontally via Kafka consumer groups. Adding pods increases
concurrency only up to the partition count of the input topic. Scaling therefore
requires topic partition planning, idempotent handlers, and a clear policy for
external side effects. Three mechanisms are available, from manual to automated.

\subsubsection{Manual scaling}

\begin{verbatim}
kubectl scale deployment/<processId>-<taskId> --replicas=4
\end{verbatim}

Suitable for development, staging, or controlled incident response. The operator
watches consumer lag, processing latency, error rate and downstream service
capacity before changing replicas.

\subsubsection{KEDA autoscaling (volume-based)}

When \texttt{--strimzi} or \texttt{--separate-workers} is active, the
scaffolder generates \texttt{keda-scalers.yml} with one
\texttt{ScaledObject} per worker. Each has dual triggers --- Kafka consumer
lag as the primary signal, and Prometheus latency as a quality-of-service
safety net:

\begin{verbatim}
apiVersion: keda.sh/v1alpha1
kind: ScaledObject
metadata:
  name: invoice_receipt-review_invoice
spec:
  scaleTargetRef:
    name: invoice_receipt-review_invoice
  minReplicaCount: 1
  maxReplicaCount: 10
  triggers:
    - type: kafka
      metadata:
        bootstrapServers: kafka-broker:9092
        consumerGroup: invoice_receipt_review_invoice_worker
        topic: invoice_receipt_review_invoice_input
        lagThreshold: "100"
    - type: prometheus
      metadata:
        serverAddress: http://durga-monitoring:8081
        metricName: durga_activity_latency_ms
        threshold: "30000"
        query: durga_activity_latency_ms{pct="p95",
          process_id="invoice_receipt",
          activity_id="review_invoice"}
\end{verbatim}

The Kafka trigger scales on message backlog (volume). The Prometheus trigger
scales when processing latency degrades despite low lag --- for example, when
a worker is bottlenecked on an external service or database.

\subsubsection{Custom metrics with Prometheus Adapter + HPA (Kubernetes-native)}

An alternative for environments without KEDA: Prometheus scrapes
\texttt{/api/metrics}, the Prometheus Adapter exposes custom metrics to
the K8s metrics API, and an HPA scales the Deployment.

\begin{verbatim}
# ServiceMonitor (prometheus-operator)
apiVersion: monitoring.coreos.com/v1
kind: ServiceMonitor
metadata:
  name: durga-monitoring
spec:
  selector:
    matchLabels: { app: durga-monitoring }
  endpoints:
    - path: /api/metrics
      interval: 15s
---
# Prometheus Adapter rule (in its ConfigMap)
- seriesQuery: 'durga_activity_latency_ms{pct="p95"}'
  name:
    matches: "durga_activity_latency_ms"
    as: "activity_latency_p95_ms"
---
# HPA
apiVersion: autoscaling/v2
kind: HorizontalPodAutoscaler
metadata:
  name: review-invoice-latency
spec:
  scaleTargetRef:
    apiVersion: apps/v1
    kind: Deployment
    name: invoice_receipt-review_invoice
  minReplicas: 1
  maxReplicas: 10
  metrics:
    - type: Pods
      pods:
        metric:
          name: activity_latency_p95_ms
        target:
          type: AverageValue
          averageValue: "30000"
\end{verbatim}

\subsubsection{Metrics reference}

The monitoring app exposes at \texttt{/api/metrics}:

\begin{description}
\item[\texttt{durga\_activity\_latency\_ms}] Gauge. Labels: \texttt{pct}
  (p50, p95, p99, avg), \texttt{process\_id}, \\ \texttt{activity\_id}.
  Value is milliseconds.
\item[\texttt{durga\_activity\_samples}] Gauge. Labels: \texttt{process\_id},
  \texttt{activity\_id}. Total completed samples.
\item[\texttt{durga\_activity\_sla\_violations}] Counter. Labels:
  \texttt{process\_id}, \texttt{activity\_id}. Count exceeding the
  configured (global) SLA-latency threshold.
\item[\texttt{durga\_alarm\_active}] Gauge (1 per active alarm). Labels:
  \texttt{process\_id}, \texttt{activity\_id}, \texttt{syndrome},
  \texttt{severity}. Every active alarm-state entry, including SLA alarms.
\item[\texttt{durga\_alarm\_fire\_count}] Counter. Same labels as
  \texttt{durga\_alarm\_active}; number of firings folded into the active alarm.
\item[\texttt{durga\_sla\_last\_observed}] Gauge. Labels: \texttt{process\_id},
  \texttt{activity\_id}, \texttt{syndrome}. Latest observed SLA measurement ---
  wall-clock latency in ms for \texttt{SLA\_LATENCY}, or windowed completion count
  for \texttt{SLA\_THROUGHPUT}.
\item[\texttt{durga\_sla\_limit}] Gauge. Same labels as
  \texttt{durga\_sla\_last\_observed}; the agreed limit --- maximum latency in ms,
  or minimum completions per window.
\item[\texttt{durga\_streams\_state}] Gauge. 1 if RUNNING, 0 otherwise.
\end{description}

\subsection{Monitoring in production}

The monitoring topology (\texttt{ProcessMonitoringApp}) is a separate
deployment that consumes the canonical \texttt{process-events-\{processId\}} topic and
materializes queryable state. In production:

\begin{itemize}
\item Deploy the monitoring app as a companion service alongside the pipeline
  (same Kafka cluster, different consumer group).
\item The Quarkus REST API and Svelte dashboard provide operational visibility ---
  instance counts, latency summaries, stuck-instance detection, trend reporting.
\item The monitoring topology uses global tables: every instance has a full
  local replica of the monitoring model. Any instance can answer all queries
  without cross-instance RPC.
\item Protect the REST API and dashboard with network policy, ingress
  authentication, or an internal-only service boundary. Process instance views
  may include business keys, correlation ids and error messages.
\item Retain the canonical lifecycle event topic for at least the longest
  monitoring recovery window you expect to support.
\end{itemize}

\subsection{Configuration reference}

\begin{longtable}{|>{\raggedright\arraybackslash}p{0.32\textwidth}|>{\raggedright\arraybackslash}p{0.18\textwidth}|>{\raggedright\arraybackslash}p{0.42\textwidth}|}
\hline
\rowcolor[gray]{0.9}\textbf{Variable} & \textbf{Default} & \textbf{Description} \\
\hline
\texttt{KAFKA\_BOOTSTRAP\_SERVERS} & \texttt{localhost:9094} & Kafka broker list \\
\texttt{HTTP\_PORT} & \texttt{8080} & Quarkus HTTP port \\
\texttt{PROFILE} & \texttt{dev} & Quarkus configuration profile \\
\texttt{START\_KAFKA} & \texttt{false} & Start Docker Compose Kafka before running \\
\texttt{DOCKER} & \texttt{false} & Build Docker image in deploy script \\
\texttt{K8S} & \texttt{false} & Apply Kubernetes manifests in deploy script \\
\hline
\end{longtable}
