\section{Generated Code Extension Strategy}
\label{sec:ExtensionStrategy}

Generated code is a starting point. The BPMN model is the primary design source; generated
code is scaffolding expected to be completed and adapted. Durga provides three strategies
for extending generated code, ordered by preference.

\subsection{Strategy 1: Use a registered plugin}

For common data pipeline operations, use one of the 15 registered plugins. Annotate a
\texttt{ServiceTask} with \texttt{camunda:plugin} and \texttt{camunda:pluginConfig}
in the BPMN model. The scaffolder generates a complete, working worker that delegates
to the plugin at runtime. No new code is required.

This covers: field remapping, filtering, type coercion, string templating, explicit field masking,
regex extraction, JSON flatten/unflatten, UUID injection, timestamp normalization,
schema validation, key-value enrichment, field-based routing, window counting, and
Kafka Connect source/sink configuration.

\subsection{Strategy 2: Implement a custom activity}

When no plugin covers the required logic, mark the BPMN activity with
\texttt{camunda:plugin="custom"}. The scaffolder generates:

\begin{itemize}
\item A \textbf{contract interface} (e.g. \texttt{MyTransformContract extends Plugin})
  that defines the API between generated and hand-written code.
\item A \textbf{delegating worker} that injects the contract via CDI (\texttt{@Inject}),
  delegates each message to \texttt{implementation.execute()}, and includes a null
  guard with dead-letter routing.
\end{itemize}

The developer writes an implementation class in a standard Java source file,
compiles and tests it with the rest of the project. The generated contract
interface and delegating worker are not modified by hand.

A build-time utility (\texttt{ModelEnricher}) scans compiled classes for
implementations of generated contracts, matches them to BPMN activities, and writes
\texttt{customImpl}, \texttt{customSource}, and \texttt{customHash} properties back
into the local BPMN model copy. On regeneration, the enriched model carries the
implementation references forward --- the scaffolder regenerates the contract and
worker but never touches the hand-written implementation class.

The BPMN model is copied into \texttt{src/main/resources/} in every generated project.
It is version-controlled alongside the custom implementation and contract interface.
Plugin-only workers can be \texttt{.gitignored} since they are fully regenerated from
the model.

\subsection{Strategy 3: Modify generated code directly}

For one-off customizations that don't fit the plugin or custom activity models,
modify the generated Java files directly. The scaffolder skips files that already
exist in \texttt{src/main/java/} on regeneration, so modifications survive.

This approach is reserved for cases where strategies 1 and 2 are insufficient.
Custom activity contracts separate generated code from hand-written logic, so
regeneration does not overwrite custom work.

\subsection{What is typically edited}

Common edits in generated projects:
\begin{itemize}
\item custom activity implementations (implementing generated contracts),
\item gateway condition expressions in the BPMN model (evaluated at runtime by the
  recursive descent parser --- no code modification needed),
\item payload handling and validation,
\item integration calls inside custom implementations,
\item operational configuration values in \texttt{application.yml}.
\end{itemize}

\subsection{What should remain generated}

Keep these close to the generated structure:
\begin{itemize}
\item topic and channel topology,
\item lifecycle event emission shape,
\item probe script set,
\item subprocess and boundary wiring,
\item monitoring-facing event semantics.
\end{itemize}

\subsection{Regeneration strategy}

When the pipeline needs structural changes (add/remove/reorder activities, change
gateway logic, adjust plugin configuration), edit the BPMN model in Camunda Modeler
and regenerate the project. For custom activities, start from the enriched model
that already carries implementation metadata.

Plugin workers are regenerated from scratch. Contract interfaces are regenerated if
the activity configuration changed. Hand-written implementation classes are never
touched. If an implementation class was removed or renamed, the scaffolder emits a
warning but does not delete the orphaned file.
