\section{Glossary}
\label{sec:Glossary}

\begin{description}
\item[BPMN] Business Process Model and Notation, the process-modeling notation used as input to Durga.
\item[Canonical lifecycle stream] The per-process \texttt{process-events-\%processId\%} topic that carries normalized process and activity events.
\item[Call activity] A BPMN element used to model invocation of another process-like unit, mapped in Durga through explicit call and reply topics.
\item[Event subprocess] A subprocess triggered by an event within an enclosing scope instead of by the main sequence flow.
\item[Execution topics] The generated Kafka topics that move work between tasks, gateways and handlers.
\item[Generated helper scripts] The scaffolded scripts used to run demos, inject inputs, complete tasks, force failures and observe output.
\item[Kafka Streams state store] The local materialized store used by the monitoring app to answer low-latency queries.
\item[Process orchestrator] The generated runtime class that detects terminal flows and emits explicit completion for the enclosing process.
\item[ProcessStateView] The repository's materialized per-instance monitoring projection.
\item[Scope cancellation] The generated runtime behavior that suppresses or redirects normal flow after an interrupting boundary or interrupting event subprocess fires.
\item[Validation mode] An opt-in mode (\texttt{--validation}) in which the whole process is generated as a complete shadow that reads the live production input on a dedicated consumer group, runs each plugin with a validation \texttt{PluginExecutionContext} (side effects suppressed), and writes to \texttt{-validation} topics, so its output is compared per task against the prior/production output (Section~\ref{sec:ValidationMode}).
\item[Shadow process] The complete validation-mode process: it mirrors production but reads through a dedicated consumer group and writes only to \texttt{-validation} topics and the per-process \texttt{process-events-\%processId\%-validation} topic, never to a real process topic.
\item[Plugin execution context] The \texttt{PluginExecutionContext} passed to \texttt{executeWithResult}; \texttt{validationMode()} tells side-effecting plugins to suppress external effects while still returning the response they would have produced.
\item[Validation events topic] The per-process \texttt{process-events-\%processId\%-validation} topic carrying a shadow's lifecycle events; the comparator's candidate source.
\item[Activity throughput] The per-activity items-processed count materialized into \texttt{process-activity-throughput} and served at \texttt{/api/processes/\{id\}/throughput}.
\item[Validation result] A \texttt{ValidationResult} record classifying one comparison as \texttt{EQUAL}, \texttt{DIFF}, \\ \texttt{PRIOR\_MISSING}, or \texttt{CANDIDATE\_ERROR}, with the located differences.
\item[Scaffolder] The BPMN-driven generator entry point, implemented by:
\begin{quote}
\filepath{org.gautelis.durga.tools.BpmnScaffolder}
\end{quote}
\item[Subprocess scope service] A generated class that manages entry, completion and scope-level lifecycle behavior for an embedded subprocess.
\end{description}
