\section{Monitoring Topology}
\label{sec:MonitoringTopology}

\subsection{Purpose}

The monitoring subsystem is a separate runtime concern from the generated process
execution graph. Generated handlers emit canonical lifecycle events; the monitoring
topology consumes those events and turns them into queryable operational views.

Execution topics show how work is routed. The monitoring topology answers:
\begin{itemize}
\item where is a specific process instance now,
\item how many instances are in a given state,
\item which activities are taking the longest,
\item which instances appear stuck,
\item what are the coarse trend counts for starts, completions, and failures,
\item which configured alarms are active and where they last fired.
\end{itemize}

\subsection{Topology structure}

The entry point is \texttt{ProcessMonitoringTopology.buildTopology()}. It
subscribes to \\ \texttt{process-events-.*} via regex, monitoring all processes
in a single instance.
It consumes events, filters null keys and values, and forks the stream into
five parallel sub-topologies plus a BPMN model cache, each with a dedicated
local state store:

\begin{description}
\item[Instance state] Aggregates events into \texttt{ProcessStateView} per instance.
Materialized in \\ \texttt{process-state-store}, output to \texttt{process-state}.
A filtered subset (active instances) is output to \texttt{process-active-state}.
\item[State counts] Regroups state views by \texttt{processId:state} key and counts
instances. Materialized in \texttt{process-state-counts-store}, output to
\texttt{process-state-counts}.
\item[Activity latency] Feeds events through a custom \texttt{ActivityLatencyProcessor}
that records \\ \texttt{ACTIVITY\_ENTERED} timestamps in a persistent key-value store
(\texttt{process-activity-entry-store}) and, on later completion or termination events,
emits duration deltas. These deltas are aggregated into rolling \texttt{ActivityLatencyStats}
per activity, then mapped to \texttt{ActivityLatencySummary} and output to
\texttt{process-latency}.
\item[Trends] Filters to \texttt{PROCESS\_STARTED}, \texttt{PROCESS\_COMPLETED},
and \texttt{PROCESS\_FAILED} events, maps them to hourly trend bucket keys,
counts per bucket, and outputs \texttt{ProcessTrendPoint} records to
\texttt{process-trends}. Materialized in \texttt{process-trends-store}.
\item[Queryable replicas] Six \texttt{builder.globalTable()} calls consume the
output topics above plus \\ \texttt{process-models}, each materialized as a global table
with a \texttt{-global-store} or \texttt{-store} suffix. These are full replicas on
every monitoring instance and are what the query service reads.
\item[BPMN model cache] A global table on \texttt{process-models} stores the latest
BPMN 2.0 XML per process ID (String key, String value). Processes publish their
models here on startup via \texttt{BpmnModelPublisher} or the generated
\texttt{ModelRegistration} bean. The diagram endpoint serves models from this cache.
\end{description}

Fault detection and alarm state are deliberately separate stream topologies.
\texttt{FaultDetectionTopology} consumes \texttt{process-events-.*} and
\texttt{process-models}, evaluates BPMN alarm configuration, and emits
\texttt{AlarmEvent} records to \texttt{fault-alarms}. \texttt{AlarmStateTopology}
then consumes \texttt{fault-alarms} and folds repeated firings into
\texttt{AlarmStateView} records in \texttt{alarm-state-store}. The monitoring
application runs these as separate \texttt{KafkaStreams} instances with separate
application IDs and state directories.

Validation mode adds a fourth separate topology. \texttt{ValidationTopology}
merges candidate task outputs (\texttt{validation-candidate-outputs}) with the
prior/production output (the \texttt{ACTIVITY\_COMPLETED} lifecycle event),
keyed on \texttt{processId:activityId:processInstanceId}, and materializes a \\
\texttt{ValidationResult} per task per instance in \texttt{validation-results-store}.
It runs as its own \texttt{KafkaStreams} instance and is described in
Section~\ref{sec:ValidationMode}.

\subsection{State stores}

The topology maintains ten state stores:

\newcommand{\storetable}[1]{%
\begin{longtable}{|>{\raggedright\arraybackslash}p{0.40\textwidth}|>{\raggedright\arraybackslash}p{0.20\textwidth}|>{\raggedright\arraybackslash}p{0.32\textwidth}|}
\hline
\rowcolor[gray]{0.9}\textbf{Store} & \textbf{Type} & \textbf{Purpose} \\
\hline
\endfirsthead
\hline
\rowcolor[gray]{0.9}\textbf{Store} & \textbf{Type} & \textbf{Purpose} \\
\hline
\endhead
#1
\end{longtable}
}

\storetable{
\texttt{process-state-store} & Local KTable aggregate & Per-instance latest \texttt{ProcessStateView} \\
\texttt{process-state-counts-store} & Local KTable count & Counts per \texttt{processId:state} key \\
\texttt{process-latency-store} & Local KTable aggregate & Rolling \texttt{ActivityLatencyStats} per activity \\
\texttt{process-activity-entry-store} & Persistent KV & Timestamps for the latency temporal join \\
\texttt{process-trends-store} & Local KTable count & Trend bucket counts \\
\texttt{process-models-store} & Global table & BPMN 2.0 XML per process ID (process registry) \\
\texttt{*-global-store} (5$\times$) & Global table & Full replicas for HTTP/CLI query access \\
}

The five local stores drive computation — aggregation, counting, and temporal joins.
The five global stores serve queries. The output topics (\texttt{process-state},
\texttt{process-state-counts}, \\ \texttt{process-active-state}, \texttt{process-latency},
\texttt{process-trends}) decouple computation from query access: any monitoring instance
can spin up and receive a full read replica without recomputing.

The fault and alarm-state stores are separate from the core process read model:

\begin{itemize}
\item \texttt{fault-counts-store}, \texttt{fault-windows-store}, and
      \texttt{fault-last-alarm-store} support fault detection syndromes and alarm
      suppression.
\item \texttt{alarm-state-store} is the queryable active-alarm read model consumed by
      the HTTP API and dashboard.
\end{itemize}

\subsection{Latency computation}

Activity latency uses a temporal join pattern without windowed stream-stream joins.
The \\ \texttt{ActivityLatencyProcessor} stores each \texttt{ACTIVITY\_ENTERED} event's
timestamp under the key \texttt{instanceId:activityId} in the persistent
\texttt{process-activity-entry-store}. When a later \texttt{ACTIVITY\_COMPLETED},
\texttt{ACTIVITY\_ESCALATED}, \texttt{ACTIVITY\_CANCELLED}, \texttt{GATEWAY\_TAKEN}, \\ 
\texttt{PROCESS\_COMPLETED}, or \texttt{PROCESS\_FAILED} event arrives for the same
key, the processor computes the duration and emits an \texttt{ActivityLatencyDelta}.

These deltas are grouped by a canonical activity key (processId, version, activityId,
event type) and aggregated into \texttt{ActivityLatencyStats} — a running count,
sum, min, and max — then mapped to \texttt{ActivityLatencySummary} for output.

\subsection{Processing guarantees}

The topology is configured with \texttt{AT\_LEAST\_ONCE} semantics. This trades
exactly-once guarantees for higher throughput and lower latency. State views may
transiently reflect duplicate or stale updates, which is acceptable for a monitoring
read model. The state directory defaults to \filepath{target/kafka-streams-state}
and can be overridden via the \texttt{durga.streams.state.dir} system property.

\subsection{Query access pattern}

The process query service (\texttt{ProcessMonitoringQueryService}) accesses only the global
table stores via \texttt{KafkaStreams.store()}. It uses point lookups for instance
state and full-store scans for counts, latency, trends, and stuck detection — filtering
in memory by process ID where needed. No interactive queries or cross-instance RPC is
required, since global tables replicate the entire content to every monitoring instance.

The alarm query service (\texttt{AlarmStateQueryService}) reads
\texttt{alarm-state-store} from the alarm-state \texttt{KafkaStreams} instance and
scans it for global, process-specific, and latest-instance-specific alarm views.

\subsection{Relationship to demos}

Repository-level demo publishers and scenario runners emit synthetic lifecycle
events directly to \\ \texttt{process-events}, making the monitoring subsystem
independently testable without a full generated process runtime.

\subsection{Current shape}

The monitoring path now supports:
\begin{itemize}
\item process self-registration: generated projects publish their BPMN models on startup,
\item process discovery from the model cache, so new processes appear before events flow,
\item queryable state replicated from monitoring topics into global read-side stores,
\item latency summaries maintained incrementally via a custom processor,
\item active-instance queries using a dedicated active-state index,
\item coarse history as bucketed trend counts for process starts, completions, and failures,
\item per-process BPMN diagrams served from the model cache with activity overlays,
\item fault detection from BPMN alarm configuration,
\item queryable active-alarm state derived from \texttt{fault-alarms},
\item per-task validation comparisons (candidate vs.\ prior output) derived from
      \texttt{validation-\\candidate-outputs}, materialized in
      \texttt{validation-results-store}.
\end{itemize}

History is intentionally coarse — Durga supports basic trend reporting but is
optimized more for current-state inspection than long-retention analytical
reporting.
