\section{Monitoring Interfaces}
\label{sec:MonitoringInterfaces}

The monitoring subsystem exposes these interfaces over Kafka Streams read models:
\begin{itemize}
\item a Quarkus HTTP API and embedded SPA,
\item a command-line client.
\end{itemize}

\subsection{HTTP service}

Entry point: \texttt{ProcessMonitoringApp} (Quarkus). Query layers:
\texttt{ProcessMonitoringQueryService} for process state and
\texttt{AlarmStateQueryService} for alarm state.

Endpoints (all under \texttt{/api/}):
\begin{itemize}
\item \texttt{/health} --- Kafka Streams readiness for probes and load balancers,
\item \texttt{/api/health} --- API-namespace health alias,
\item \texttt{/api/processes/list} --- all known process IDs (from counts + model registry),
\item \texttt{/api/instances/\{processInstanceId\}} --- latest instance view,
\item \texttt{/api/processes/\{processId\}/counts} --- process-specific state counts,
\item \texttt{/api/processes/\{processId\}/latency} --- activity latency summaries,
\item \texttt{/api/processes/\{processId\}/trends} --- bucketed start/completion/failure history,
\item \texttt{/api/counts} --- global state-count view,
\item \texttt{/api/alarms} --- global active alarm-state view,
\item \texttt{/api/processes/\{processId\}/alarms} --- process-specific active alarm-state view,
\item \texttt{/api/instances/\{processInstanceId\}/alarms} --- active alarms whose latest firing belongs to the instance,
\item \texttt{/api/stuck} --- stuck-instance detection,
\item \texttt{/api/validation/summary?processId=\textit{id}} --- per-task validation outcome counts (see Section~\ref{sec:ValidationMode}),
\item \texttt{/api/validation/results?processId=\&taskId=\&status=} --- individual validation comparisons,
\item \texttt{/api/validation/instances/\{processInstanceId\}} --- validation comparisons for one instance,
\item \texttt{/api/diagram?processId=\textit{id}} --- BPMN 2.0 XML from the model cache,
\item \texttt{/api/metrics} --- Prometheus-format metrics.
\end{itemize}

The diagram endpoint serves models from the \texttt{process-models-store}
Kafka state store, where processes self-register by publishing their BPMN
XML on startup.

The HTTP API is designed for internal operational use. Put it behind the same
network and identity controls used for other production admin interfaces before
exposing it outside a trusted environment.

\subsection{CLI client}

The repository includes a convenience script at \filepath{durga-monitor}:

\begin{verbatim}
./durga-monitor health
./durga-monitor counts invoice_receipt
./durga-monitor latency invoice_receipt
./durga-monitor stuck invoice_receipt 300
./durga-monitor instance <processInstanceId>
\end{verbatim}

Defaults to \texttt{http://localhost:8081}. Override with
\texttt{DURGA\_URL}. Equivalent to:

\begin{verbatim}
java -cp target/durga-0.1.0-beta.1.jar \
  org.gautelis.durga.monitoring.ProcessMonitoringClient \
  http://localhost:8081 <command>
\end{verbatim}

\subsection{Dashboard}

The Svelte SPA is built from \filepath{monitoring-ui/} and served directly
from the monitoring backend. It shows:
\begin{itemize}
\item a process inventory with all known process IDs and aggregate KPIs,
\item alarm KPIs and process-level alarm counts,
\item per-process drill-down: state mix, lifecycle trends, activity latency, stuck instances, and alarm state,
\item a validation report: per-task outcome counts and per-instance input/prior/candidate diffs when candidate tasks run in validation mode (see Section~\ref{sec:ValidationMode}),
\item instance detail lookup with latest alarm details,
\item a BPMN diagram viewer with color-coded activity overlays (alarms, latency, and state).
\end{itemize}

The diagram viewer uses \texttt{bpmn-js} with built-in zoom (scroll wheel),
pan (grab tool), and fit-to-viewport controls.

The monitoring SPA is served from the same Quarkus HTTP server that hosts the
REST API. It polls \texttt{/api/*} endpoints and renders an interactive
dashboard with a BPMN diagram viewer.

\subsection{How the interfaces fit together}

The interfaces consume the same materialized read models:
\begin{enumerate}
\item processes self-register by publishing BPMN models to \texttt{process-models},
\item lifecycle events flow through \texttt{process-events-*},
\item the fault detector emits matching \texttt{AlarmEvent} records to \texttt{fault-alarms},
\item Kafka Streams materializes state, counts, latency, trends, the model cache, and alarm state,
\item the HTTP layer exposes that materialized state,
\item the CLI and dashboard consume the same HTTP API.
\end{enumerate}

The operator workflow is:
\begin{itemize}
\item deploy a process --- it auto-registers its model on boot,
\item the dashboard shows the new process in the inventory,
\item publish a scenario or task outcome,
\item watch lifecycle events,
\item query one instance or inspect counts, latency, and alarms,
\item view the BPMN diagram with live overlays.
\end{itemize}
