\section{Monitoring and Observability}
\label{sec:Monitoring}

\subsection{Canonical process-events topic}

Durga emits normalized lifecycle events on the logical channel \texttt{process-events}.
Generated workers, gateways, subprocess handlers, boundary handlers and helper publishers
all emit events describing process progress, waiting states, completion, failure,
escalation and cancellation. The default physical topic is
\texttt{process-events-\{processId\}} so that each process definition is isolated at the
transport level. The monitoring topology subscribes to \texttt{process-events-.*} to
consume events from all processes.

This separates execution transport from queryable state. Execution topics remain useful for
routing, but the monitoring model depends on projections built from explicit lifecycle events.
For production use, treat the lifecycle topic as an operational contract: retention,
access control, schema compatibility and replay policy determine how far back monitoring can
recover and how safely downstream tools can consume process state.

\subsection{Process self-registration}

Processes self-register with the monitoring server by publishing their BPMN 2.0 model
to the \texttt{process-\\models} Kafka topic. Each generated project includes a
\texttt{ModelRegistration} CDI bean that runs at startup:

\begin{itemize}
\item reads the BPMN file from the classpath (\texttt{src/main/resources/\{name\}.bpmn}),
\item publishes it to \texttt{process-models} with the process ID as the Kafka message key,
\item logs success or failure without blocking the process startup.
\end{itemize}

The monitoring server caches models in the \texttt{process-models-store} state store. \\
The \texttt{/api/diagram?processId=\textit{id}} endpoint serves the cached model.
Model updates overwrite previous versions (latest-wins semantics).

This registry makes processes discoverable: \texttt{/api/processes/list} merges
process IDs from both the count store (processes with live instances) and the model
store (registered processes), so newly deployed processes appear in the dashboard
before their first lifecycle event arrives.

\subsection{Kafka Streams topology}

The monitoring topology subscribes to \texttt{process-events-.*} via a regex
pattern, monitoring all processes in a single instance. State stores use
unsuffixed names. The SPA shows an inventory of all known processes with
drill-down to per-process detail.

The topology consumes lifecycle events and materializes:
\begin{itemize}
\item latest process state by instance,
\item counts by current state,
\item latency summaries,
\item stuck-instance views,
\item trend buckets,
\item BPMN model cache.
\end{itemize}

The main entry point is \texttt{org.gautelis.durga.monitoring.ProcessMonitoringApp}.

\subsection{HTTP API}

The monitoring service exposes endpoints for health, process discovery, instance
lookup, counts, latency, stuck instances, alarm state, diagram serving, and a dashboard.
All monitoring paths are served under the \texttt{/api/} prefix; a root \texttt{/health} endpoint is also available.

The local development loop is:
\begin{enumerate}
\item generate a process,
\item the process auto-registers its BPMN model on boot,
\item publish scenarios or task inputs,
\item observe lifecycle events,
\item query current state and latency,
\item view the BPMN diagram with activity overlays,
\item iterate on the BPMN model or generated runtime.
\end{enumerate}

In production, the loop is stricter: monitor Streams state, query-store
readiness, consumer lag on lifecycle and monitoring topics, stuck-instance
counts, latency percentiles, DLQ volume and failed lifecycle events. The
dashboard is an operator view over the Kafka Streams read model, not a
replacement for Kafka broker, consumer group and infrastructure monitoring.

\subsection{Fault detection}

\label{sec:FaultDetection}

The fault detection topology consumes the same \texttt{process-events-.*}
stream as the monitoring topology and evaluates each lifecycle event against
a set of configured \texttt{AlarmConfig} rules. When a rule's criteria are
met, an \texttt{AlarmEvent} is published to the \texttt{fault-alarms} topic
for downstream notification, alerting, or dashboard integration.

\subsubsection{Alarm layering}

Alarms originate from three provenance layers, identified by
\texttt{AlarmOrigin}. When multiple configs match the same scope the higher layer
takes precedence.

\begin{description}
\item[Automatic] Monitor-owned, always active, requiring no process configuration.
  Includes built-in stall and cascade detection (see STUCK / CASCADE below).
  Tunable via system properties \\ (\texttt{durga.alarm.stuck.timeoutSeconds} etc.).
\item[Opt-in] A monitor-provided default that a process enables and may tune by
  including the alarm's properties in its BPMN model.
\item[Explicit] Fully specified by a process, typically through BPMN extension
  properties \\ (\texttt{durga:alarm:<id>:<field>}) or authored directly in the
  Camunda Modeler plugin.
\end{description}

\subsubsection{Alarm configuration}

A single \texttt{AlarmConfig} record combines scope, a syndrome, severity, and a
message template:

\begin{description}
\item[Scope] Each config can target a specific process (\texttt{processId}),
  a specific activity (\texttt{activityId}), both, or neither (all events).
  A null \texttt{processId} matches every process; a null \texttt{activityId}
  matches process-level aggregates.
\item[Event type] Which \texttt{ProcessEvent.EventType} triggers evaluation
  (e.g. \texttt{PROCESS\_FAILED}, \\ \texttt{ACTIVITY\_ESCALATED}). Optional for
  STUCK and CASCADE, which are evaluated by a wall-clock punctuator rather than
  per incoming event.
\item[Syndrome] How observed behaviour is aggregated into an alarm (see below).
\item[Severity] \texttt{WARN} or \texttt{CRITICAL}.
\item[Threshold] Maximum count before alarm (required for COUNTED, SLIDING\_WINDOW,
  CASCADE; ignored for HARD\_ERROR and STUCK).
\item[Window duration] Sliding-window size for SLIDING\_WINDOW / CASCADE, or
  idle-timeout for STUCK.
\item[Message] A template rendered at alarm time; supports
  \texttt{\$\{processId\}}, \texttt{\$\{activityId\}},
  \texttt{\$\{count\}}, \texttt{\$\{processInstanceId\}}, and
  \texttt{\$\{idleSeconds\}} (STUCK only).
\item[Origin] One of \texttt{AUTOMATIC}, \texttt{OPT\_IN}, or \texttt{EXPLICIT}.
\end{description}

\subsubsection{Syndromes}

\paragraph{Event-driven syndromes}

The three event-driven syndromes are evaluated for each incoming lifecycle
event.

\begin{description}

\item[HARD\_ERROR] First occurrence triggers alarm immediately. No
threshold---every matching event produces an alarm. Suitable for critical
failures that must never happen (e.g. schema validation failure on a
compliance-critical field).

\item[COUNTED] Counts occurrences per (config, processId, activityId).
The alarm fires once when the accumulated count exceeds
\texttt{threshold}, then suppresses until the count drops back below the
threshold (re-arm). Suitable for transient errors that become significant
in volume (e.g. 5 task failures per process).

\item[SLIDING\_WINDOW] Counts occurrences within a sliding time window.
Old occurrences leak out after \texttt{windowDuration}. The alarm fires
once when the windowed count exceeds \texttt{threshold}, then suppresses
until the count drops back below threshold (re-arm). Suitable for
burst detection (e.g. more than 3 \texttt{PROCESS\_FAILED} events in a
60-second window).

\end{description}

\paragraph{Absent-progress syndromes}

The two absent-progress syndromes react to the \emph{absence} of lifecycle
events rather than events that arrive. They are evaluated by a wall-clock
punctuator scanning the active-instance index, not by the event-driven loop.

\begin{description}

\item[STUCK] An active instance that has emitted no lifecycle event for
longer than \texttt{windowDuration} (the idle-timeout) is reported as stuck.
The alarm fires once per instance; subsequent punctuation cycles suppress
until a fresh lifecycle event re-arms tracking. The template variable
\texttt{\$\{idleSeconds\}} reports how long the instance has been idle.

\item[CASCADE] Collapses a surge of STUCK alarms into a single higher-severity
alert. When the number of instances that newly became stuck within a rolling
\texttt{windowDuration} exceeds \texttt{threshold}, one CASCADE alarm fires.
Suppressed until the window count falls back below threshold (re-arm).
A CASCADE alarm's \texttt{processId} is \texttt{*} (system-wide); per-process
CASCADE configs carry the actual process identifier.

\end{description}

\subsubsection{Alarm emission}

Triggered alarms are serialized as JSON and published to the
\texttt{fault-alarms} topic. The topic is both an integration point for
external notification systems and the input to Durga's alarm-state read model.
Each \texttt{AlarmEvent} carries:

\begin{itemize}
\item \texttt{alarmId} --- unique per firing,
\item \texttt{configId} --- back-reference to the triggering config,
\item \texttt{syndrome}, \texttt{severity}, \texttt{message} --- from config,
\item \texttt{processId}, \texttt{processInstanceId}, \texttt{activityId} --- scope,
\item \texttt{triggerEventType}, \texttt{triggerTimestamp} --- what fired it
  (\texttt{triggerEventType} is \texttt{null} for CASCADE alarms),
\item \texttt{count} --- current accumulated count at firing,
\item \texttt{threshold} --- config threshold (0 for HARD\_ERROR, timeout-seconds for STUCK).
\end{itemize}

\subsubsection{Alarm state read model}

The monitoring server also starts an alarm-state topology that consumes
\texttt{fault-alarms} and materializes one \texttt{AlarmStateView} per
(processId, activityId, configId) scope in \texttt{alarm-state-store}. Process-wide
alarms use \texttt{*} as the activity key component. Repeated alarm firings are
folded into the same view:

\begin{itemize}
\item \texttt{status} is currently \texttt{ACTIVE};
\item \texttt{severity} keeps the strongest observed severity;
\item \texttt{firstTriggeredAt} and \texttt{lastTriggeredAt} delimit the firing range;
\item \texttt{fireCount} counts folded alarm firings;
\item \texttt{lastProcessInstanceId}, \texttt{lastCount}, \texttt{lastAlarmId}, and
      \texttt{lastMessage} describe the newest firing.
\end{itemize}

The REST API exposes this read model through \texttt{/api/alarms},
\texttt{/api/processes/\{processId\}/alarms}, and
\texttt{/api/instances/\{processInstanceId\}/alarms}. The dashboard uses these
endpoints for the alarm KPI, process inventory alarm counts, alarm-state table,
instance alarm details, and BPMN diagram alarm overlays.
System-wide alarms (\texttt{processId:\,*}) are excluded from the process
inventory and rendered as \texttt{system-wide} in the Alarm State widget.

\subsubsection{State stores}

The fault detection topology maintains five persistent state stores:

\begin{itemize}
\item \texttt{fault-config-store} --- serialized \texttt{AlarmConfig} per
  processId, updated when a process model is published to
  \texttt{process-models}.
\item \texttt{fault-counts-store} --- current count per (config, processId, activityId).
\item \texttt{fault-windows-store} --- list of epoch-millis timestamps per scope
  (SLIDING\_WINDOW and CASCADE).
\item \texttt{fault-last-alarm-store} --- count at which the last alarm fired,
  used for single-fire suppression.
\item \texttt{fault-seen-store} --- per-instance record of the last lifecycle
  activity observed (processId, activityId, last-seen epoch millis, alarm
  flag), used by the STUCK punctuator.
\end{itemize}

\subsubsection{Example configuration}

\begin{verbatim}
AlarmConfig hard = new AlarmConfig("schema-fail-critical", null,
    "validate_high_value", ProcessEvent.EventType.ACTIVITY_ESCALATED,
    AlarmSyndrome.HARD_ERROR, 0, null,
    AlarmSeverity.CRITICAL, "Schema validation escalated",
    AlarmOrigin.EXPLICIT);

AlarmConfig counted = new AlarmConfig("proc-fail-warn", "order-proc",
    null, ProcessEvent.EventType.PROCESS_FAILED,
    AlarmSyndrome.COUNTED, 5, null,
    AlarmSeverity.WARN, "${count} failures in ${processId}",
    AlarmOrigin.EXPLICIT);

AlarmConfig window = new AlarmConfig("burst-detect", null, null,
    ProcessEvent.EventType.PROCESS_FAILED,
    AlarmSyndrome.SLIDING_WINDOW, 3, Duration.ofMinutes(1),
    AlarmSeverity.CRITICAL, "Burst: ${count} failures in 1 min",
    AlarmOrigin.EXPLICIT);

AlarmConfig stuck = new AlarmConfig("builtin:stuck", null, null, null,
    AlarmSyndrome.STUCK, 0, Duration.ofSeconds(120),
    AlarmSeverity.WARN, "Stuck ${processInstanceId} in ${activityId} "
        + "for ${idleSeconds}s",
    AlarmOrigin.AUTOMATIC);

AlarmConfig cascade = new AlarmConfig("builtin:cascade", null, null, null,
    AlarmSyndrome.CASCADE, 5, Duration.ofSeconds(60),
    AlarmSeverity.CRITICAL,
    "Cascade: ${count} instances stalled within 60s (threshold ${threshold})",
    AlarmOrigin.AUTOMATIC);
\end{verbatim}

\subsubsection{Integration with the monitoring server}

The monitoring server starts the fault detection topology alongside the
monitoring read-model topology as a separate \texttt{KafkaStreams} instance
with its own application ID and state directory. Both subscribe to
\texttt{process-events-.*}; the fault detector also consumes
\texttt{process-models} so BPMN alarm properties can update runtime alarm
configuration. A third \texttt{KafkaStreams} instance maintains the alarm-state
read model from \texttt{fault-alarms}. Health output reports all three stream
states: core monitoring, fault detection, and alarm state.

Unit tests are in \texttt{FaultDetectionTest}, \texttt{BpmnAlarmConfigParserTest},
\texttt{AlarmStateViewTest},
\texttt{AlarmStateTopologyTest}, and \texttt{PluginResultTest};
see Appendix~\ref{sec:IntegrationTestAppendix} for the test catalogue.
