\section{Plugin Architecture}
\label{sec:PluginArchitecture}

Durga supports a plugin-driven data pipeline model where BPMN \texttt{ServiceTask}
elements can be annotated with \texttt{camunda:plugin} attributes that resolve to
registered, governed data processing components.

\subsection{Plugin registry}

Plugins are defined by YAML descriptors under \filepath{plugins/} and indexed by
\filepath{plugins/catalog.yml}. Each descriptor declares:

\begin{itemize}
\item \texttt{id} --- unique plugin identifier (used in \texttt{camunda:plugin}),
\item \texttt{name} / \texttt{version} --- human-readable identifier,
\item \texttt{category} --- governed taxonomy (transform, validate, enrich, route, aggregate, inspect, store, connect),
\item \texttt{status} --- lifecycle state (experimental, stable, deprecated, retired),
\item \texttt{implementation} --- Java class implementing
\filepath{org.gautelis.durga.plugins.Plugin}. For the Rust code-generation
target (Section~\ref{sec:RustTarget}) the same plugin is implemented natively in
the \filepath{durga-rust} crate under the same catalog identity and configuration
schema; there are no adapters onto the Java implementation.
\item \texttt{config} --- typed configuration schema (field types, defaults, allowed values).
\end{itemize}

\subsection{Governed categories}

Each plugin category has an intended contract for topic naming, lifecycle events,
and error channel conventions. The generated plugin executor now emits
category-specific lifecycle events for the supported governed cases:
\texttt{GATEWAY\_TAKEN} for route plugins, \texttt{ACTIVITY\_ESCALATED} for
validate plugin failures, and, for aggregate plugins that keep a window open by
returning no output, \texttt{PROCESS\_COMPLETED} for the absorbed instance (it is
folded into the window and reaches a terminal state rather than lingering as
active). Other plugin categories retain the normal activity completion/failure
lifecycle.

\begin{longtable}{|>{\raggedright\arraybackslash}p{0.22\textwidth}|>{\raggedright\arraybackslash}p{0.70\textwidth}|}
\hline
\rowcolor[gray]{0.9}\textbf{Category} & \textbf{Contract} \\
\hline
\textbf{transform} & 1:1 input$\to$output, ACTIVITY\_COMPLETED, dlq on failure \\
\textbf{validate} & Output + invalid branch, ACTIVITY\_ESCALATED on invalid \\
\textbf{enrich} & Input$\to$enriched output, may block on external lookup \\
\textbf{route} & Input$\to$N downstream channels, GATEWAY\_TAKEN lifecycle \\
\textbf{aggregate} & Windowed counting, ACTIVITY\_ENTERED on window start \\
\textbf{inspect} & Payload inspection and metadata extraction without changing the data-plane artifact \\
\textbf{store} & Payload externalization and retrieval through \texttt{DataHandle} references \\
\textbf{connect} & Kafka Connect source/sink configuration generation \\
\hline
\end{longtable}

\subsection{Plugin interface}

All 16 processing plugins implement \filepath{org.gautelis.durga.plugins.Plugin},
which provides two \texttt{default} methods. The two connect catalog descriptors
(\texttt{kafka-connect-source} and \\ \texttt{kafka-connect-sink}) are code-generation
descriptors that reference the scaffolder, not runtime \texttt{Plugin}
implementations.

\begin{quote}
\texttt{byte[] execute(byte[] payload, String config) throws Exception} \\
\texttt{String execute(String payload, String config) throws Exception}
\end{quote}

The generated worker calls the \texttt{byte[]} overload. Text / JSON plugins
override the \texttt{String} variant instead, with UTF-8 conversion handled by
the \texttt{byte[]} default. Binary plugins override the \texttt{byte[]} variant
directly. Only one of the two needs to be overridden.

Generated executor classes instantiate the plugin at field level and delegate
through \\ \texttt{PluginExecutionSupport.execute(...)} for each message. The config
string comes from the BPMN \texttt{camunda:pluginConfig} property.
Default execution is unchanged: the current \\ \texttt{ProcessEvent.payload} JSON is
passed to the plugin. Handle-aware execution can also be requested:

\begin{itemize}
\item \texttt{handleMode=manual} passes the \texttt{DataHandle} JSON to the plugin
      so the plugin can operate on the handle itself.
\item \texttt{handleMode=materialize} reads the bytes referenced by
      \texttt{dataHandle.uri}, passes those bytes to the plugin, writes plugin
      output back to the configured object store, and forwards a new
      \texttt{DataHandle}.
\item \texttt{pluginConfig=...} supplies the plugin's own configuration when the
      outer config string is used for handle-control options.
\end{itemize}

The returned bytes are decoded as UTF-8. JSON object results replace the
\texttt{ProcessEvent.payload}. JSON scalar, JSON array, and non-JSON text results
are wrapped as \texttt{\{ "\_value": ... \}}. A \texttt{null} result means the
incoming payload is forwarded unchanged. Generated plugin and custom workers also
emit a Vannak-compatible \texttt{DataIndividualMetadataEvent} to the
\texttt{vannak-metadata-events} topic after successful execution.

\subsection{Plugin catalog}

The repository ships with 18 catalog descriptors organized by category
(16 processing plugins implementing \texttt{Plugin} plus 2 connect code-generation
descriptors). Stable plugins are
fully supported; experimental plugins generate with a warning.

\subsubsection{Transform plugins (9)}

\begin{description}
\item[JSON Transform] Dot-notation field remapping with colon-syntax renaming and literal injection.
\item[Field Filter] Whitelist/blacklist field filtering with optional flatten.
\item[Type Coercion] Coerces field values between string, int, long, double, decimal, and boolean.
\item[String Template] Renders a template string with \texttt{\$\{field\}} substitution. Dot-notation access for nested fields.
\item[Mask] Masks explicitly configured text fields with configurable mask character and boundary character preservation.
\item[Regex Extract] Extracts named capture groups from a source field into the payload. Supports multiple matches.
\item[JSON Flatten] Flattens nested JSON to dot-notation keys, or unflattens dot-notation back to nested objects. Configurable separator and max depth.
\item[UUID Inject] Injects UUIDs (v7, v4 or v1) into specified payload fields.
\item[Timestamp Normalize] Converts between epoch\_s, epoch\_ms, ISO 8601, RFC 3339, and custom DateTimeFormatter patterns. Configurable timezone.
\end{description}

\subsubsection{Validate plugins (1)}

\begin{description}
\item[JSON Schema Validator] Validates payloads against a JSON Schema subset. Routes invalid messages to DLQ, skipped, or fails. Status: experimental.
\end{description}

\subsubsection{Enrich plugins (1)}

\begin{description}
\item[KV Enricher] Looks up a key field value in an inline data map and shallow-merges enrichment data. Status: experimental.
\end{description}

\subsubsection{Route plugins (1)}

\begin{description}
\item[Field Router] Routes messages to named output channels based on a field value. Status: experimental.
\end{description}

\subsubsection{Aggregate plugins (1)}

\begin{description}
\item[Window Counter] Counts messages within a tumbling time window and emits a summary record on window close. Optional group-by field. Status: experimental.
\end{description}

\subsubsection{Inspect plugins (1)}

\begin{description}
\item[Format Detector] Detects MIME type, textual/binary shape, JSON object/array/scalar form, byte count, and checksums. Status: experimental.
\end{description}

\subsubsection{Store plugins (2)}

\begin{description}
\item[Object Store Collector] Writes the current payload to a local file-backed object store and emits a \texttt{DataHandle}-style JSON reference. Status: experimental.
\item[Object Store Extractor] Resolves a \texttt{DataHandle}-style object reference and emits the referenced bytes. Status: experimental.
\end{description}

\subsubsection{Connect plugins (2)}

\begin{description}
\item[Kafka Connect Source] Generates Kafka Connect source connector configuration for pipeline intake topics.
\item[Kafka Connect Sink] Generates Kafka Connect sink connector configuration for pipeline egress.
\end{description}

\subsection{BPMN binding}

Plugin-annotated tasks use Camunda extension properties on \texttt{ServiceTask}
elements:

\begin{verbatim}
<bpmn:serviceTask id="transform" name="Transform Data">
  <bpmn:extensionElements>
    <camunda:properties>
      <camunda:property name="plugin" value="json-transform" />
      <camunda:property name="pluginConfig" value="name, email" />
    </camunda:properties>
  </bpmn:extensionElements>
</bpmn:serviceTask>
\end{verbatim}

Without a \texttt{camunda:plugin} attribute, \texttt{ServiceTask} elements fall
back to the existing generic worker generation.

\subsection{Generated executor}

The scaffolder resolves the plugin descriptor, validates it against the registry,
and generates a concrete executor class using the \texttt{pluginExecutorClass}
template. The executor:

\begin{itemize}
\item imports the plugin implementation class and the \texttt{Plugin} interface,
\item imports \texttt{PluginExecutionSupport} for handle-aware execution,
\item instantiates the plugin as a private field,
\item delegates each incoming message to \texttt{PluginExecutionSupport.execute(...)},
\item emits \texttt{ACTIVITY\_COMPLETED} on success,
\item emits a companion \texttt{DataIndividualMetadataEvent} to \texttt{vannak-metadata-events} on success,
\item routes failures to a dead-letter topic (\texttt{\%task\%\_dlq}) with structured error records,
\item respects \texttt{ScopeCancellationRegistry} for interrupting boundary support.
\end{itemize}

\subsection{Custom activities}

When no registered plugin covers a transformation, the \texttt{camunda:plugin="custom"}
marker triggers a different code generation path designed for hand-written implementations:

\begin{verbatim}
<bpmn:serviceTask id="custom_transform" name="Custom Transform">
  <bpmn:extensionElements>
    <camunda:properties>
      <camunda:property name="plugin" value="custom" />
      <camunda:property name="pluginConfig" 
                        value="org.example.CustomTransformContract" />
    </camunda:properties>
  </bpmn:extensionElements>
</bpmn:serviceTask>
\end{verbatim}

Instead of generating a self-contained worker, the scaffolder produces:

\begin{enumerate}
\item A \textbf{contract interface} (\texttt{CustomTransformContract}) extending \texttt{Plugin} --- this defines the API between generated and hand-written code.
\item A \textbf{delegating worker} (\texttt{CustomTransformWorkerService}) that injects the contract via CDI (\texttt{@Inject}) and delegates each message through \texttt{PluginExecutionSupport.execute(...)}. The worker includes a null guard that routes to the dead-letter queue if no implementation is present.
\end{enumerate}

The developer writes an implementation class (\texttt{CustomTransformImpl implements \\ CustomTransformContract}) in a standard Java source file. It is compiled and tested with the rest of the project. The generated contract and worker are not modified by hand.

\subsubsection{Round-trip model enrichment}

A build-time utility (\texttt{ModelEnricher}) scans compiled classes for implementations
of generated contracts, matches them to BPMN activities by contract name, and writes
\texttt{customImpl}, \texttt{customSource}, and \texttt{customHash} Camunda extension
properties back into the local copy of the BPMN model. Detection uses a
\texttt{URLClassLoader} and reflection (\texttt{isAssignableFrom}); hashing uses
the source file (\texttt{.java}) rather than the bytecode (\texttt{.class}), so the
hash is stable across JDK versions and compiler flags. The source directory is
derived from the classes directory by Maven convention (\texttt{target/classes}
$\to$ \texttt{src/main/java}) or passed explicitly.

The enriched model becomes the source of truth for future regenerations. When the
pipeline changes, regeneration starts from the enriched model --- it already knows
which activities have custom implementations and which Java class backs each one.
The custom implementation files are never touched by the scaffolder.

\subsection{Current limitations}

\begin{itemize}
\item Plugin descriptors are read from the filesystem or classpath at scaffold time --- there is no runtime plugin discovery or hot-reload.
\item Config validation at scaffold time is not yet implemented --- invalid configs are caught at runtime by the generated executor.
\item The plugin catalog must be maintained manually; there is no registry publishing or version-resolution tooling.
\end{itemize}
