\section{Testing And Verification}
\label{sec:TestingVerification}

\subsection{Generator verification}

Three layers:
\begin{itemize}
\item dry-run tests that treat scaffold summaries and previews as a contract,
\item generation tests that write projects to temp directories and verify output
  (file existence, plugin imports, condition expressions, contract interfaces),
\item a compilation integration test that generates a full project from
  \filepath{invoice\_receipt.bpmn} and compiles all \texttt{.java} files with
  \texttt{javax.tools.JavaCompiler}, verifying the generated code is
  syntactically and semantically correct.
\end{itemize}

The verification set spans 24 BPMN fixtures covering:
\begin{itemize}
\item message and signal events,
\item timers and boundaries,
\item subprocesses and nested subprocesses,
\item call activities,
\item event subprocesses (message, signal, timer, error, escalation),
\item data pipelines (plugin-annotated tasks, custom activities),
\item gateway condition expressions.
\end{itemize}

\subsection{Plugin verification}

Each of the 16 Java plugin implementations has dedicated unit tests
covering normal operation, edge cases, and error handling. The plugin
registry loading and descriptor validation are exercised by the scaffolder
tests. Dedicated tests also cover format detection, object-store collection and
extraction, handle-aware plugin execution, and Vannak metadata event creation.

\subsection{Data pipeline verification}

Four dedicated tests verify:
\begin{itemize}
\item the plugin-annotated pipeline (\filepath{data\_pipeline\_demo.bpmn})
  generates correct plugin executor classes and data-asset metadata,
\item the order events pipeline (\filepath{order\_events\_pipeline.bpmn})
  generates workers for 8 plugin types and an XOR gateway with
  runtime-evaluated conditions,
\item the log processing pipeline (\filepath{log\_processing\_pipeline.bpmn})
  generates workers for regex extraction, templating, flattening, and
  validation plugins,
\item the custom activity demo (\filepath{custom\_activity\_demo.bpmn})
  generates contract interfaces and delegating workers.
\end{itemize}

\subsection{Connect verification}

Connect tests verify that \texttt{--connect} generates generic source and sink
connector configs with correct topic lists for both dry-run and full generation
modes. A dedicated data-store test verifies that BPMN data stores generate
per-store connector skeletons under \filepath{connect/data-stores/}, including
S3, Neo4j and PostgreSQL sink examples.

\subsection{Model enricher verification}

Three tests verify the \texttt{ModelEnricher}: enrichment of custom
activity metadata, no-op when no implementation is found, and property
update on re-enrichment. Tests compile real Java source files during
execution and verify that \texttt{customSource} references the
\texttt{.java} path (not the \texttt{.class} file) and that hashes are
computed from source files.

\subsection{Core runtime contract verification}

\begin{itemize}
\item \texttt{ProcessEvent} serialization round-trips and compact-constructor inference,
\item \texttt{ProcessState} serialization round-trips,
\item \texttt{ScopeCancellationRegistry} cancellation tracking, merging, clearing and null-guard behavior,
\item \texttt{ProcessStateView} projection tests covering cancellation, escalation,
process completion, gateway events, concurrent activities, and state-key generation.
\end{itemize}

\subsection{Monitoring verification}

\begin{itemize}
\item activity latency statistics (min, max, avg, p50, p95, p99, SLA violations),
\item demo publishers and scenario runners that drive the monitoring topology
without a full generated process runtime,
\item fault detection unit tests (14 scenarios) covering the three syndromes
(HARD\_ERROR, COUNTED, SLIDING\_WINDOW), config scoping, event type
filtering, alarm suppression, and re-arm behaviour (see
Section~\ref{sec:FaultDetection}).
\item alarm state unit tests covering state-key generation, repeated firing
fold-up, severity preservation, late older events, and topology construction.
\end{itemize}

\subsection{CI pipeline}

The non-Docker verification command \texttt{mvn test -Dtest='!*IntegrationTest'}
currently runs 260 tests. Docker-backed integration tests are suffixed
\texttt{IntegrationTest} and are run separately for release gates.

\subsection{Integration tests}

Five integration test suites use Testcontainers to spin up real Kafka brokers
and verify runtime behaviour end to end. Each suite is documented in detail in
Appendix~\ref{sec:IntegrationTestAppendix}.

\begin{description}
\item[ProcessEventKafkaIntegrationTest] Serialization and JSON round-trip
integrity over a live Kafka broker. Verifies that \texttt{ProcessEvent}
payloads, status codes, error info, and compact-constructor inference survive
wire transport intact.

\item[ProcessStateStoreIntegrationTest] Compacted-topic state persistence via
\texttt{ProcessStateStore}. Exercises save/load, latest-wins semantics,
multiple keys, not-found, and timeout handling against a real compacted
Kafka topic.

\item[MonitoringTopologyIntegrationTest] The monitoring topology end to end:
state materialization, state counts, activity latency summaries, stuck
instance detection, and unkeyed event handling.

\item[ChaosIntegrationTest] Topology resilience under restart: verifies state
recovery after crash, absence of duplicate state entries, and three
consecutive restarts without corruption.

\item[E2EPipelineIntegrationTest] Full data pipeline end-to-end test. Verifies
the scaffolder generates category-specific lifecycle events
(\texttt{GATEWAY\_TAKEN} for route plugins, \texttt{ACTIVITY\_ESCALATED} for
validate plugins, \texttt{ACTIVITY\_ENTERED} for aggregate plugins), and that
the monitoring topology correctly materializes state and latency from these
event types in a mixed-type lifecycle.
\end{description}

\subsection{Automation}

The script \filepath{scripts/run-e2e-test.sh} automates the full test cycle:

\begin{verbatim}
./scripts/run-e2e-test.sh test       # Integration tests (Testcontainers)
./scripts/run-e2e-test.sh feed       # Infra + monitor + data feed
./scripts/run-e2e-test.sh test-run   # Tests + monitor + continuous feed
./scripts/run-e2e-test.sh stop       # Tear down
\end{verbatim}

The \texttt{feed} and \texttt{test-run} commands build and start the
monitoring server (port 8081), launch a Redpanda broker, PostgreSQL, and
Kafka Connect via Docker Compose, then inject the \texttt{E2ETestDataLoader}
which publishes variable order records for live observation.

\subsection{Not fully covered}

\begin{itemize}
\item load and concurrency behavior,
\item multi-instance monitoring scenarios.
\end{itemize}
